Online {\IterComp}

### 2017-11-15 results.tex
rest of results

Figure~\ref{fig:error-O3} shows that profiling with relaxation incurs very little dynamic error. It also confirms that the \WCRelaxLower
relaxation is overly conservative, while the \WPRelaxLower relaxation achieves a better trade-off. Although the error bound is not
guaranteed by the \WPRelaxLower relaxation, its dynamic error is always below the 5\% threshold. For both relaxation strategies the 2\%
threshold is too conservative to be useful. The rest of the paper only evaluates the 5\% threshold.

@@

In this section we evaluate the effectiveness of our work efficiency metric.
We perform online {\itercomp} using all five profiling strategies, namely
\OracleRM, \OraclePP,\OptProf,\WCRelax-5\%, and \WPRelax-5\%.
Because the \OracleRM measures actual speedup, it provides reference results for
assessing the effectiveness of the work-efficiency metric.

For all strategies, the same optimization sequence is used for multiple inputs, using a dynamic input-window size, as explained in
Section~\ref{sec:oic-infra}. The average performance over the input window provides an estimate for the overall performance of the
optimization sequence across distinct inputs. The optimization sequences are ranked based on their average performance, and the best
optimization sequence is selected.

The comparison between \OracleRM and \OraclePP is crucial for validating the
effectiveness of the work efficiency metric in guiding online {\itercomp},
while the other configurations demonstrate the viability of applying online
{\itercomp} in real-world scenarios.

In order to evaluate the quality of the final optimization sequences selected by each configuration of the online {\itercomp}, we compare
their speedup over the \texttt{-O3} optimization across all the 1,000 inputs of the benchmark being optimized. None of the configurations
degrades the performance over \texttt{-O3} (with statistical significance).
%When measuring the wall-clock time for each input, to reduce noise, we execute
%the same input until we have a statistically sound measurement, i.e. we execute
%until we have an interval no larger than 1\% with 99\% confidence.
Figure~\ref{fig:speedups} shows these average speedups for all test benchmarks.
This figure shows that the best optimization sequence selected with the Oracle-PP
is very close to the performance of the best optimization sequence selected with
the \OracleRM, where \OraclePP achieves on average about 80\% of the performance
improvement obtained by the \OracleRM.
This result is important for demonstrating that our work efficiency metric has
the potential to produce good results in real-world online scenarios.

The \OptProf achieves 4\% improvement on average, which represents 45\% of the
performance improvement obtained by the \OracleRM.
This difference to \OraclePP is explained by the real (intrusive) profiling used
in the \OptProf configuration.
The profiling affects the search in two key ways:
\textit{(i.)} the overhead incurred by the profiling affects the execution time,
and therefore also affects the work efficiency metric,
\textit{(ii.)} the instrumentation code may affect many of the optimizations, e.g.,
due to the use of a global variable or in decisions taken based on a cost-model
for the instructions.
However, even in the \OptProf configuration, which has a highly intrusive instrumentation,
it is still possible to obtain performance improvements in a realistic online
{\itercomp}.

Finally, the evaluation also indicates that the use of relaxation algorithms
is beneficial not only for its reduction in overhead, which is directly observed
by the user, by it also tends to improve the quality of the search.
This improvement has to do with the aforementioned effects of an intrusive
instrumentation on the search.
The relaxation has the potential to reduce both effects of the profiling.
Both the \WCRelax-5\% and \WPRelax-5\% achieve about 60\% of the
performance improvement obtained by the \OracleRM.

>>>>

Figure~\ref{fig:error-O3} shows that profiling with relaxation incurs very little dynamic error. It also confirms that the \WCRelaxLower
relaxation is overly conservative, while the \WPRelaxLower relaxation achieves a better trade-off. Although the error bound is not
guaranteed by the \WPRelaxLower relaxation, its dynamic error is always below 5\%. For both relaxation strategies the 2\% threshold is too
conservative to be useful. The rest of the paper only evaluates the 5\% threshold.

@@

In this section we evaluate the effectiveness of our work efficiency metric. We perform online {\itercomp} using all five profiling
strategies, namely \OracleRM, \OraclePP, \OptProf, \WCRelax-5\%, and \WPRelax-5\%. We evaluate every optimization sequence on multiple
inputs, using a dynamic input-window size, as explained in Section~\ref{sec:oic-infra}. We use the average work efficiency over the input
window to rank different optimization sequences and we select the one with the highest work efficiency.

\OracleRM makes iterative compilation decisions based on the actual speedup over the unoptimized version. It does not depend on any kind of
instrumentation or work estimation, so it acts as the baseline against which the rest of the profiling strategies are compared. \OraclePP
is also not affected by instrumentation but uses work efficiency instead of speedup. We use it to decide whether our work metric can be
used for iterative compilation on principle. The other configurations demonstrate the viability of applying online {\itercomp} in
real-world scenarios.

We evaluate the quality of the selected optimization sequences by comparing their average speedup over \texttt{-O3} across all 1,000 inputs
of each benchmark.
% None of the configurations degrades the performance over \texttt{-O3} (with statistical significance). \FIXME{PP: According to the figure there are slowdowns!!!}
Figure~\ref{fig:speedups} shows these speedups for all test benchmarks. In most cases, \OraclePP is very close to \OracleRM, achieving on
average 80\% of its speedup. This demonstrates that it is a valid choice to drive iterative compilation decision using work efficiency. The
few cases where the two strategies produce noticeably different results are for the \texttt{tiff} benchmarks. The weaker correlation between
work and unoptimized runtime for these benchmarks is to blame for the lower speedups achieved by \OraclePP.

The \OptProf achieves 4\% improvement on average, which represents a little bit more that half of the speedup of \OraclePP. The culprit is
the one thing that is different between the two strategies, the profiling instrumentation. It affects the search in two key ways:
\textit{(i.)} the overhead incurred by profiling affects the execution time and, through it, the work efficiency metric,
\textit{(ii.)} the inserted instrumentation code occasionally changes how optimizations are applied, e.g., due to the use of a global
variable or in decisions taken based on a cost-model for the instructions. With the \OptProf configuration, instrumentation is very
intrusive and this translates into lower speedups or even slowdowns for five benchmarks. Just measuring work with traditional profiling
approaches is not enough for online iterative compilation.

This is where our two relaxation algorithms become particularly important. Beyond just reducing the user perceived overhead of iterative
compilation, they also improve the search itself by keeping the behavior of the instrumented binary as close as possible to that of the
uninstrumented one. Relaxation reduces the deleterious effects of instrumentation on both the runtime and how optimizations are applied.
As a result, \WCRelax-5\% and \WPRelax-5\% are very close to the behavior of \OraclePP and they achieve about 60\% of the performance
improvement obtained by \OracleRM.
